In this appendix, we compute the one-loop corrections to the light neutrino masses in our dark neutrino model, following Refs.~\cite{Kniehl:1996bd,Grimus:2002nk,AristizabalSierra:2011mn}.

We work with the on-shell (OS) renormalization scheme. This is ensured by requiring that the off-diagonal elements of the self-energy be diagonal when the external particles are on their mass shell, and that the residue of the renormalized propagator are equal to one. Note this is only applicable to the off-diagonal entries that involve at least one heavy neutrino, and that the light-light entries are all non-zero and finite at one-loop.

Assuming Majorana neutrino fields, one can write the self-energy tensor in its most general form:
%
\begin{equation}
  \Sigma_{ij} (\slashed{q}) = \slashed{q} {\rm P}_L \Sigma_{ij}^L (q^2) + \slashed{q} {\rm P}_R \Sigma_{ij}^R (q^2) +  {\rm P}_L \Sigma_{ij}^M (q^2) +  {\rm P}_R \Sigma_{ij}^{M*} (q^2),
\end{equation}
%
where by virtue of the Majorana nature the previous terms obey
\[\Sigma_{ij}^L (q^2) = \Sigma_{ij}^{R*} (q^2), \qquad \Sigma_{ij}^M (q^2) = \Sigma_{ji}^M (q^2). \]

\subsection{Self-energy}

For now, we will ignore kinetic and scalar mixing effects, as these can be show to be small given current experimental bounds. The contribution from the scalar fields $s= h^0, \varphi^0$, the goldstones $G = G_h, G_\varphi$ and the vector bosons $V = Z, Z^\prime$ are 

\begin{align*}
 -i\,\Sigma^s_{ij} (p^2) &= (-i)^2 \left(\Delta_s P_R + \Delta^*_s P_L \right)_{ik} \times\\&\qquad\qquad \int \frac{d^d k}{(2\pi)^d} \frac{i (\slashed{p} + \slashed{k} + m_k) }{(p+k)^2 - m_k^2} \frac{i}{k^2 - m_s^2} \left( \Delta_s P_R + \Delta^*_s P_L \right)_{kj},\\
 %
 -i\,\Sigma^G_{ij} (p^2) &= (-i)^2 \left(i\,\Delta_G P_R + i\,\Delta^*_G P_L \right)_{ik} \times\\&\qquad\qquad\int \frac{d^d k}{(2\pi)^d} \frac{i (\slashed{p} + \slashed{k} + m_k) }{(p+k)^2 - m_k^2} \frac{i}{k^2 - \xi_V m_V^2} \left( i\,\Delta_G P_R + i\,\Delta^*_G P_L  \right)_{kj},\\
%
 -i\,\Sigma^G_{ij} (p^2) &= -(-i)^2 \gamma^\mu \left( C_V P_L - C_V^T P_R \right)_{ik} \times \\&\qquad\qquad\int \frac{d^d k}{(2\pi)^d} \frac{i (\slashed{p} + \slashed{k} + m_k) }{(p+k)^2 - m_k^2} \frac{iP_{\mu\nu}}{k^2 - m_V^2} \gamma^\nu \left( C P_L - C^T P_R \right)_{kj},
\end{align*}
with no index summation notation. In the latter term, we defined the vector boson propagator numerator, which we rewrite as
%
\begin{align*}
\gamma^\mu P_{\mu\nu} \gamma^\nu &= \gamma^\mu \left[ g_{\mu\nu} - (1-\xi_V) \frac{k_\mu k_\nu}{k^2 - \xi_V m_V^2} \right] \gamma^\nu\\
&= d - (1-\xi_V) \frac{k^2 - m_k^2}{k^2 - \xi_V m_V^2} - \frac{m_k^2}{m_V^2} \frac{ (k^2 - \xi_V m_V^2) - (k^2 - m_V^2)}{k^2 - \xi_V m_V^2}.
\end{align*}

This allows us to write the relevant part of the self-energy as functions of the scalar two-point loop function
\begin{align}
 B_0 (l, m_a^2,m_c^2) = \mu^{2\epsilon} \int \frac{d^d k}{(2 \pi)^d} \frac{1}{(k^2-m_a^2)( (l+k)^2 - m_c^2 )},
\end{align}
such that
\begin{align}
  \Sigma^s_{ij} (0) \, P_R =& -\frac{\pi^2}{(2\pi)^4} \mu^{d-4} \left[ (\Delta_s)_{ik} m_k B_0 (0,m_k^2, m_s^2) (\Delta_s)_{kj}\right] \, P_R \\
%
 \Sigma^G_{ij} (0) \, P_R =& \frac{\pi^2}{(2\pi)^4} \mu^{d-4} \left[ (\Delta_G)_{ik} m_k B_0 (0,m_k^2, \xi_V m_V^2) (\Delta_G)_{kj}\right] \, P_R \\
%
 \Sigma^V_{ij} (0) \, P_R =& -\frac{\pi^2}{(2 \pi)^4} \mu^{d-4} \left[ (C_V)_{ik} m_k \,f( m_k^2, m_V^2, \xi m_V^2) (C_V^*)_{kj} \right] \, P_R,
\end{align}
where the rearrangement of the boson propagator allowed us to write $f( m_k^2, m_V^2, \xi m_V^2)$ as
\begin{align*}
f( m_k^2, m_V^2, \xi m_V^2) =&\, d\, B_0(0,m_k^2, m_V^2) - (1-\xi_V) B_0 (0,m_V^2, \xi_V m_V^2) +\\ & \frac{m_k^2}{m_V^2} B_0 (0, m_k^2, \xi_V m_V^2) - \frac{m_k^2}{m_V^2} B_0 (0,  m_k^2, m_V^2). 
\end{align*}

Finally, the scalar loop function is given by
\begin{align*}
 B_0 (0,m_a^2,m_b^2) &= \frac{1}{\epsilon} - \gamma_E + \ln{4 \pi} - \int_0^1 dx \ln{\frac{m_a^2-x(m_a^2 - m_b^2)}{\mu^2}}\\
 &= \frac{1}{\epsilon} - \gamma_E + \ln{4 \pi} - \frac{m_a^2}{m_b^2 - m_a^2} \left[ \ln{\frac{m_a^2}{\mu^2}}  - 1 \right] + \frac{m_b^2}{m_b^2 - m_a^2} \left[ \ln{\frac{m_b^2}{\mu^2}}  - 1 \right].
\end{align*}

The finiteness of our final result and its gauge invariance are a consequence of the two following identities
\begin{align}
 \Delta_G = \Delta_S = C \frac{\hat{m}}{m_V} + \frac{\hat{m}}{m_V}C^T, \quad \quad C \hat{m} C^T = 0.
\end{align}

For light neutrinos ($i,j=1,2,3$), the final result reads

\begin{align}
 \Sigma_{ij} P_R  = -\frac{\pi^2}{(2\pi)^4} C \hat{m}\left[ d\, B_0(0,\hat{m}^2,m_V^2) + \frac{\hat{m}^2}{m_V^2} \left( B_0 (0,\hat{m}^2,m_S^2) - B_0(0,\hat{m}^2,m_V^2) \right) \right] C^T P_R.
\end{align}

After significant algebra, and checking the cancellation of divergencies, we arrive at 
%
\begin{align}\label{eq:masses_general}
 m_{ij} = \frac{1}{4\pi^2}\sum_{k=4}^5 \Big[ & C_{ik} C_{jk} \frac{m_k^3}{m_Z^2}F(m_k^2,m_Z^2,m_h^2)  \,+\, D_{ik} D_{jk} \frac{m_k^3}{m_{Z^\prime}^2}F(m_k^2,m_{Z^\prime}^2,m_{\varphi^\prime}^2) \Big],
\end{align} 
%
where
%
\begin{equation} \label{eq:loop_function}
F(a,b,c) \equiv \frac{3 \, \ln{(a/b)}}{a/b - 1}  + \frac{\ln{(a/c)}}{a/c - 1}.
\end{equation}
%

For the full expression, involving mixing, we find ($B_0(x^2) \equiv B_0(0,\hat{m}^2,x^2)$) \textcolor{red}{needs some checking! This is only for light neutrinos}
%
\begin{align*}
%
    \Sigma^M(0) &= C\hat{m}\left\{ 4B_0(m_Z^2) -
\frac{\hat{m}^2}{m_Z^2}\left[B_0(m_Z^2) - c_{\omega^\prime}^2B_0(m_{h^\prime}^2)- s_{\omega^\prime}^2B_0(m_{\varphi^\prime}^2)\right] \right\}C^\text{T}\\
%
&~+D\hat{m}\left\{ 4B_0(m_{Z^\prime}^2) -
\frac{\hat{m}^2}{m_{Z^\prime}^2}\left[B_0(m_{Z^\prime}^2) - c_{\omega^\prime}^2B_0(m_{\varphi^\prime}^2)- s_{\omega^\prime}^2B_0(m_{h^\prime}^2)\right] \right\}D^\text{T}\\
%
&~+s_{\omega^\prime} c_{\omega^\prime} \left\{C\hat{m}\left[B_0(m_{h^\prime}^2) -B_0(m_{\varphi^\prime}^2) \right]D^\text{T}+D\hat{m}\left[ B_0(m_{h^\prime}^2) -B_0(m_{\varphi^\prime}^2) \right]C^\text{T}\right\}.
%
\end{align*}
